Sufficient Conditions for Reachability in Automata Networks with Priorities

### 2015-06-05 parts-example.tex
proofread + move contexts to appendix

At this point, the static analysis results presented in
\pref{ssec:ua} can be used to check if the model is functional,
\ie if gene $a$ can oscillate, thus leading to the formation of stripes.
Starting from context $\ctx^1 = \PHstate{f_1, a_0, c_0}$,
we thus want to check the reachability of $\myp^1 = a_1$;
then, starting from any new state obtained,
we want to check the reachability of $\myp^2 = a_0$,
and finally $\myp^3 = a_1$ once again to ensure that we entered a cycle.
%This consists in using 
For this, we apply the method proposed in \pref{ssec:ordered-ua}:
we consider each reachability step independently
and we use \pref{th:approxinf} three times;
the initial states of steps 2 and 3 are computed with the
extraction method of \pref{ssec:concret}.

We first build the local causality graph $\thisB{\myp^1}{\ctx^1}$
related to the reachability of $\myp^1 = a_1$ from the initial state $\ctx^1$.
This graph is depicted in \pref{fig:metazoan-sa}(left).
\pref{th:approxinf} allows to conclude that this reachability is true.
One can thus extract a concretizing scenario from this local causality graph
with a traversal of the graph, as explained in \pref{ssec:concret}.
Such a traversal is detailed in \pref{tab:concret-metazoan};
the scenario extracted from this example consists of only one action:
$\PHfrappes{c_0, f_1}{a_0}{a_1}$.
Another traversal of the same graph
would consist in visiting node $f_1$ before node $c_0$
when first descending from node $\{ c_0, f_1 \}$;
however, the same scenario would be extracted.
As there exists no other traversal, we thus have in this case:
$\Delta(\thisB{\myp^1}{\ctx^1}) = \{ \PHfrappes{c_0, f_1}{a_0}{a_1} \}$.

We denote in the following:
$\ctx^2 = \PHstate{f_1, a_0, c_0} \play \PHfrappes{c_0, f_1}{a_0}{a_1} =
\PHstate{f_1, a_1, c_0}$
the state resulting from the play of the scenario concretizing
$\myp^1$ in $\ctx^1$.
As explained in \pref{ssec:ordered-ua},
one can use this resulting state in order to carry on with another
successive reachability,
such as the reachability of $\myp^2 = a_0$ in our case.
The local causality graph used to check the reachability of $\myp^2$
from $\ctx^2$ is given in \pref{fig:metazoan-sa}(middle).
The same reasoning allows to conclude that this reachability is true, and
to extract the following set containing only one concretizing scenario:
$\Delta(\thisB{\myp^2}{\ctx^2}) =
\{ \PHfrappes{a_1, f_1}{c_0}{c_1} \concat \PHfrappes{c_1}{a_1}{a_0} \}$.
We note that once again, two traversals are possible
(by visiting node $a_1$ before or after note $f_1$)
but both output the same scenario, and therefore end up in the same sate
$\ctx^3 = \PHstate{f_1, a_0, c_1}$.
Finally, the last local causality graph $\thisB{\myp^3}{\ctx^3}$,
depicted in \pref{fig:metazoan-sa}(right),
allows to conclude that this final reachability is true.

@@

    (tc10s) edge (tnc10s)
    (tnc10s) edge (ta0)
    (ta0) edge (ta00)
    (ta00) edge (ta00s)
    (ta00s) edge (tna00s)
    ;
  \end{tikzpicture}
  \caption{%
  \label{fig:metazoan-sa}%
    The three successive saturated graphs of local causality
    of the AAN in \pref{fig:metazoan-ph}
    for the successive reachability of $a_0$, $a_1$ and $a_0$
    from the initial context
    $\ctx = \PHstate{f_1, a_0, c_0}$.
    The (left) graph allows to check the reachability of $a_1$
    from the initial context $\ctx$.
    The (middle) graph is for the reachability of $a_0$
    and the context $\PHstate{f_1, a_1, c_0}$.
    The (right) graph is for the last reachability, $a_1$,
    and the context $\PHstate{f_1, a_0, c_1}$.
  }
\end{figure}

>>>>

At this point, the static analysis results presented in
\pref{ssec:ua} can be used to check if the model is functional,
\ie if gene $a$ can oscillate, thus leading to the formation of stripes.
Starting from state $s^1 = \PHstate{f_1, a_0, c_0}$,
we thus want to check the reachability of $\myp^1 = a_1$;
then, starting from any new state obtained,
we want to check the reachability of $\myp^2 = a_0$,
and finally $\myp^3 = a_1$ once again to ensure that we entered a cycle.
%This consists in using 
For this, we apply the method proposed in \pref{ssec:ordered-ua}:
we consider each reachability step independently
and we use \pref{th:approxinf} three times;
the initial states of steps 2 and 3 are computed with the
extraction method of \pref{ssec:concret}.

We first build the local causality graph $\thisB{\myp^1}{s^1}$
related to the reachability of $\myp^1 = a_1$ from the initial state $s^1$.
This graph is depicted in \pref{fig:metazoan-sa}(left).
\pref{th:approxinf} allows to conclude that this reachability is true.
One can thus extract a concretizing scenario from this local causality graph
with a traversal of the graph, as explained in \pref{ssec:concret}.
Such a traversal is detailed in \pref{tab:concret-metazoan};
the scenario extracted from this example consists of only one action:
$\PHfrappes{c_0, f_1}{a_0}{a_1}$.
Another traversal of the same graph
would consist in visiting node $f_1$ before node $c_0$
when first descending from node $\{ c_0, f_1 \}$;
however, the same scenario would be extracted.
As there exists no other traversal, we thus have in this case:
$\Delta(\thisB{\myp^1}{s^1}) = \{ \PHfrappes{c_0, f_1}{a_0}{a_1} \}$.

We denote in the following:
$s^2 = \PHstate{f_1, a_0, c_0} \play \PHfrappes{c_0, f_1}{a_0}{a_1} =
\PHstate{f_1, a_1, c_0}$
the state resulting from the play of the scenario concretizing
$\myp^1$ in $s^1$.
As explained in \pref{ssec:ordered-ua},
one can use this resulting state in order to carry on with another
successive reachability,
such as the reachability of $\myp^2 = a_0$ in our case.
The local causality graph used to check the reachability of $\myp^2$
from $s^2$ is given in \pref{fig:metazoan-sa}(middle).
The same reasoning allows to conclude that this reachability is true, and
to extract the following set containing only one concretizing scenario:
$\Delta(\thisB{\myp^2}{s^2}) =
\{ \PHfrappes{a_1, f_1}{c_0}{c_1} \concat \PHfrappes{c_1}{a_1}{a_0} \}$.
We note that once again, two traversals are possible
(by visiting node $a_1$ before or after note $f_1$)
but both output the same scenario, and therefore end up in the same sate
$s^3 = \PHstate{f_1, a_0, c_1}$.
Finally, the last local causality graph $\thisB{\myp^3}{s^3}$,
depicted in \pref{fig:metazoan-sa}(right),
allows to conclude that this final reachability is true.

@@

    (tc10s) edge (tnc10s)
    (tnc10s) edge (ta0)
    (ta0) edge (ta00)
    (ta00) edge (ta00s)
    (ta00s) edge (tna00s)
    ;
  \end{tikzpicture}
  \caption{%
  \label{fig:metazoan-sa}%
    The three successive saturated graphs of local causality
    of the AAN in \pref{fig:metazoan-ph}
    for the successive reachability of $a_0$, $a_1$ and $a_0$
    from the initial state
    $s = \PHstate{f_1, a_0, c_0}$.
    The (left) graph allows to check the reachability of $a_1$
    from the initial state $s$.
    The (middle) graph is for the reachability of $a_0$
    and the state $\PHstate{f_1, a_1, c_0}$.
    The (right) graph is for the last reachability, $a_1$,
    and the state $\PHstate{f_1, a_0, c_1}$.
  }
\end{figure}

### 2015-06-05 parts-sa.tex
proofread + move contexts to appendix

We present a static analysis that takes as input an AAN
$\PH = (\PHs; \PHl; \PHa)$,
an initial state, and a local state of an automaton in $\PHs$, for identifying sufficient conditions
that ensure the existence of a scenario starting from the initial state and leading to a state
where the given local state is present.

@@

An objective sequence can be seen as an abstract representation of a set of scenarios that describe
(part of) the successive state changes of the automata.
We denote by $\concr(\w)$ (\pref{def:concr}) the set of scenarios matching with an objective sequence
$\w$ in the state $s\in\PHs$.
It is essentially all the scenarios for which there exists a mapping from the bounces of each
objective to the bounce of an action in the scenario which preserve the sequential ordering.

@@

\modmf{%
The previous subsections permitted to define all the tools needed to build the Local Causality Graph (LCG).
This graph is denoted $\cwB$ where $u$ is the local state to be reached
from the initial context $\ctx$.
and an example is depicted in \pref{fig:sa-livelock}.
It will be the basis of our static analysis,
as it represents the causality links between the different objectives involved
in the solving of the reachability of $u$.
This LCG is built recursively by considering some required local states (\eg $u$),
linking them to objectives (\eg $\PHobj{w}{u}$, if $w \in \ctx$),
and locally refining these objectives in order to include new required local states
from other automata.}

Technically,
one can iteratively refine a given objective into (possibly several) objective
sequences following its bounce sequences.
Indeed, the refinement of an objective $P$ with $\zeta\in\BS(P)$
consists in prepending to $P$ the
objectives leading to the activation of all hitters of the actions in $\zeta$,
in the same sequential order.
The generalization of this refinement to an objective sequence would
naturally require to consider all possible
interleaving between the refinements~\cite{PMR12-MSCS}.

@@

Formally, a Local Causality Graph (LCG)
is a digraph where $\cwV \subseteq \Proc\cup \Obj \cup \Sol \cup \sSol$ is the set of vertices,
with $\sSol=\powerset(\Proc)$ and $\Sol=\Obj\times\powerset(\sSol)$,
and $\cwE \subseteq \cwV \times \cwV$ is the set of oriented edges
(\pref{def:lcg}).
% 
% These refinements can be summarized with a so-called Local Causality Graph (LCG).
% Given an initial context $\ctx$ and a local state $\myp\in\Proc$,
% an LCG $\cwB = (\cwV, \cwE)$
% is a digraph where $\cwV \subseteq \Proc\cup \Obj \cup \Sol \cup \sSol$ is the set of vertices
% where $\sSol=\powerset(\Proc)$ and $\Sol=\Obj\times\powerset(\sSol)$
% and $\cwE \subseteq \cwV \times \cwV$ is the set of oriented edges
% (\pref{def:lcg}).
A node in $\cwV\cap\Obj$ represents an objective to refine.
Such a node is linked to nodes in $\Sol$ which represent
sets of hitter sets extracted from a
bounce sequence of the objective
(that is, one node for each bounce sequence in $\BS(P)$); thus,
given an objective $P$, if $\zeta\in\BS(P)$
and $\abstr\zeta$ is the set of all hitters of $\zeta$,
% and $\abstr\zeta \DEF \{\PHhitter(\zeta_n) \mid
% n\in\indexes\zeta \}$,
then $P$ is linked to a node $\langle P, \abstr\zeta\rangle\in\Sol$
(\pref{def:lcg}\nobreakdash-\ref{lcg-obj-sol}).
For each hitter set $ps\in\abstr\zeta$,
the node $\langle P,\abstr\zeta\rangle\in\cwV\cap\Sol$ is
linked to the node $ps\in\sSol$
(\pref{def:lcg}\nobreakdash-\ref{lcg-sol-sync}).
Then, a node $ps\in\cwV\cap\sSol$ is linked to each local state $a_i \in ps$ is contains
(\pref{def:lcg}\nobreakdash-\ref{lcg-sync-ls}).
A local state node $a_i\in\cwV\cap\Proc$
is linked to each objective $\obj{a_j}{a_i} \in \Obj$, where
$a_j$ is either in the initial context $\ctx$ or
is a local state node in the LCG
(\pref{def:lcg}\nobreakdash-\ref{lcg-ls-obj}).
Finally, given an objective node $\obj{a_j}{a_i}\in\cwV\cap\Obj$,
if a local state $a_k\neq a_i$ is in its successors (which is given by
$\conn_{(\cwBNodes,\cwBEdges)}(\obj{a_j}{a_i})$), then
$\obj{a_j}{a_i}$ is linked to $\obj{a_k}{a_i}$ in the LCG
(\pref{def:lcg}\nobreakdash-\ref{lcg-conn}).

\begin{definition}
\label{def:lcg}
Given a context $\ctx$ and a local state $\myp\in \Proc$,
the Local Causality Graph (LCG) for reachability under-approximation
$\cwB \DEF (\Bv, \Be)$
with
$\cwBNodes \subseteq \Proc \cup \Obj \cup \NSol \cup \sSol$
and
$\cwBEdges \subseteq \cwBNodes \times \cwBNodes$,
is the smallest graph such that:
\begin{enumerate}
\item
$\myp \in \cwBNodes$
\item
$a_i\in\cwBNodes\cap\Proc \Leftrightarrow \{ (a_i,\obj{a_j}{a_i}) \mid
a_j \neq \myp \wedge (a_j\in\ctx \vee a_j\in\cwBNodes\cap\Proc)
\}\subseteq\cwBEdges$
\label{lcg-ls-obj}
\item
$P\in\cwBNodes\cap\Obj \Leftrightarrow 
	\{ (P,\langle P,\abstr \zeta\rangle) \mid \zeta\in\BS(P) \}\subseteq\cwBEdges$
\label{lcg-obj-sol}
\item
$\langle P, pps \rangle\in\cwBNodes\cap\NSol \Leftrightarrow
	\{ (\langle P,pps\rangle, ps) \mid ps\in pps \}\subseteq\cwBEdges$
\label{lcg-sol-sync}
\item
$ps \in\cwBNodes\cap\sSol \Leftrightarrow
	\{ (ps, a_i) \mid a_i\in ps \}\subseteq\cwBEdges$
\label{lcg-sync-ls}
\item
$\obj{a_i}{a_j}\in \cwBNodes\cap\Obj \Rightarrow 
	\{(\obj{a_i}{a_j},\obj{a_k}{a_j}) \mid a_k\neq a_j,$
\label{lcg-conn}
\\
\hspace*{4cm}
$a_k\in \conn_{(\cwBNodes,\cwBEdges)}(\langle\obj{a_i}{a_j},\abstr\zeta\rangle),$
\\
\hspace*{4cm}
$\zeta\in\BS(\obj{a_i}{a_j}) \}\subseteq\cwBEdges$
\end{enumerate}
with $\abstr\zeta \DEF \{\PHhitter(\zeta_n) \mid n\in\indexes\zeta \}$.
\end{definition}

\modmf{%
The definition of an LCG tackles two cases
where the bounce local state of an objective may be changed.
First,
the addition of objectives based on the local states already mentioned anywhere in the LCG,
as performed in \pref{def:lcg}\nobreakdash-\ref{lcg-ls-obj},
ensures to take into account the possible changes in the active local states
made by other objectives.
We thus try to ensure that a required local state is reachable even when
starting from a local state of the same automaton that is not in the initial context,
but that may be active at some point of the solving.
Second, \pref{def:lcg}\nobreakdash-\ref{lcg-conn},
allows to re-target objectives whose own solving changes their target local state.
This happens when playing the actions that activate the local states required to solve
an objective $P$ also changes the active local state of $\PHsort(P)$.
In this case, the initial objective $P$ is re-targeted to another
objective $\PHobj{p}{\bounce{P}}$, where $p$ is the new active local state in $\PHsort(P)$.
Linking to all objectives of this kind ensures that all possible disturbances
are taken into account.
}

\modmf{%
The LCG of \pref{def:lcg} contains synchronizing nodes ($\sSol$)
that allow to express the need for several local states simultaneously
in order to play a given action.
This is the main difference regarding~\cite{PMR12-MSCS}
which tackled with Process Hitting, whose actions are limited to at most one hitter
and thus did not require this kind of synchronization.
However, \pref{def:lcg} is equivalent to the definition of the structure
given in~\cite{FPMR13-CS2Bio},
although it was adapted for AANs which notably allowed to simplify its formalism.
}

@@

Given a context $\ctx$ and a local state $\myp\in\Proc$, the LCG $\cwB$ contains a set of objectives
that can be used to build a concrete scenario reaching $\myp$.
Because we are focused on sufficient conditions for reachability, we do not require that all
possible scenarios can be derived from the LCG.

In this section, we prove that if the LCG has no cycle, all its nodes in $\Obj$ have at least one
child, and all its nodes in $\sSol$ satisfy a
particular criteria, so-called independence, then from any state
delimited by the context $\ctx$, there exists a scenario that reaches $\myp$
(\pref{th:approxinf}).

@@

The intuition of the under-approximation of \pref{th:approxinf} is the following.
Given a state $s\subseteq\ctx$, we recursively refine the initial objective
(which is: reaching $\myp$ from the initial state) according to its children.
As the LCG is acyclic, by hypothesis, such a recursion always terminates, and as all the objective nodes have at
least one child, it never gets stuck.
The refinement of an objective node $\obj{a_i}{a_j}$ acts as follows:
if the node has another objective node $\obj{a_k}{a_j}$ as child, we first refine the objective
$\obj{a_i}{a_k}$ (which, by construction, is necessarily in the LCG)
and then we refine the objective $\obj{a_k}{a_j}$.
If the objective node has only successors in $\Sol$ (bounce sequences), one is picked arbitrarily.
If $\langle P, pps\rangle$ is the chosen node,
by construction there exists $\zeta\in\BS(P)$ such that $\abstr\zeta = pps \in \powerset(\sSol)$.
If $\zeta = \emptyseq$ (for instance in the case where $a_i = a_j$), the recursion stops and we
continue to the next stage.
Otherwise, for each $n\in\indexes\zeta$,
for each $b_i \in\PHhitter(\zeta_n)$,
we refine the objective $\obj{b_j}{b_i}$, where $b_j$ is the state of $b$ in the current state.
By induction, $\obj{b_j}{b_i}$ is a child of $b_i$ in the LCG of \pref{def:lcg}.
Thus, we know that the current state of $b$ is $b_i$.
After having repeated this procedure for each $b_i\in\PHhitter(\zeta_n)$,
because all the nodes in $\sSol$, and in particular $\PHhitter(\zeta_n)$, are independent,
we know that all the local states in $\PHhitter(\zeta_n)$ are in the current state.
In addition, we know that the state of automaton $a$ has remained unchanged, otherwise
$\obj{a_i}{a_j}$ would have an objective child.
Hence, the action $\zeta_n$ is playable in the current state.
We can thus apply this action, modifying the state of $a$ to $a_j$ and continue to the next stage.
In the end, this recursive procedures builds a scenario from $s\subseteq\ctx$ to a state containing $\myp$.

\begin{theorem}[Under-approximation]
\label{th:approxinf}
  Given an AAN $(\PHs; \PHl; \PHa)$,
  a context $\ctx$ and an objective sequence $\w$,
  if the LCG $\cwB$ contains no cycle,
  all nodes in $\Obj$ have at least one child,
  and all nodes in $\sSol$ are independent,
  then for any state $s\in\PHl$ such that $s\subseteq\ctx$,
  there exists a scenario $\delta\in\Sce$ with $\myp\in s\play\delta$.
\end{theorem}

@@

We note furthermore that the method does not require any completeness of the bounce sequences
$\BS$.
Therefore, in order to reduce the number of bounce sequences to consider, and potentially remove
cycles and non-satisfying nodes,
one can consider only a sub-set of bounce sequences
(and in particular: a single bounce sequence) for each objective.
However, such an approach requires to enumerate all possible combinations of bounce sequences
subsets, hence being exponential in the number of considered bounce sequences.
%objectives solved by at least two bounce sequences.

@@

    (b0) edge (b10) edge (b00)
    (b10) edge (b10s)
    (b10s) edge (nb10s)
    (b00) edge (b00s)
    (b00s) edge (nb00s)
    (b1) edge (b01) edge (b11)
    (b01) edge (b01s)
    (b01s) edge (b01ss)
    (b01ss) edge (a0)
    (b11) edge (b11s)
    (b11s) edge (nb11s)
    ;
    \end{tikzpicture}
  \caption{
  \label{fig:sa-livelock}
    The local causality graph $\cwB$ on the AAN in \pref{fig:ph-livelock}
    for the reachability of $\myp = c_1$
    from the initial context $\ctx = \PHstate{a_1, b_0, c_0}$.
    Elements in $\Proc$ are represented by rectangular nodes,
    elements in $\Sol$ are represented by small circles,
    and elements in $\sSol$ and $\Obj$ are the remaining borderless nodes.
    \pref{th:approxinf} is inconclusive on this example as
    node $\{ a_1, b_1 \} \in \sSol$
    is not independent (see \pref{def:coherent}).
    Indeed, $a_0$ is a successor of $b_1$, but $a_0 \neq a_1$
    (and the same also stands for $b_0$, which is a successor of $a_1$).
  }
\end{figure}
\end{example}

@@

This section gives a recursive method to find a scenario that concretizes
the reachability of a given local state $\myp \in \Proc$,
from a given initial state $s_0 \in \PHl$,
provided that \pref{th:approxinf} answered positively
for a context $\ctx \in \Ctx$ so that $s_0 \subset \ctx$.
The algorithm proposed in what follows relies on a traversal of the Local Causality Graph
$\cwB$ that has been used for this conclusion.
The correctness of this extraction can be demonstrated
in the same fashion as \pref{th:approxinf}.

@@

The execution of such a traversal outputs a scenario from $s_0$.
In the following, we denote by $\Delta(\cwB)$ the set of all scenarios extracted from the
local causality graph $\cwB$.
Indeed, several scenarios may exist for a same graph,
generated by the arbitrary choices that exist in this algorithm.
An example of such a traversal is given on a small example
in \pref{ssec:ex-metazoan}.

>>>>

We present a static analysis that takes as input an AAN
$\PH = (\PHs; \PHl; \PHa)$,
an initial state in $\PHl$, and a local state in $\Proc$.
Its objective it to identify sufficient conditions
that ensure the existence of a scenario starting from the initial state,
and leading to a state where the given local state is present.

@@

An objective sequence can be seen as an abstract representation of a set of scenarios that describe
(part of) the successive state changes of the automata.
We denote by $\concr(\w)$ (\pref{def:concr}) the set of scenarios matching with an objective sequence
$\w$ in the state $s\in\PHl$.
It is essentially all the scenarios for which there exists a mapping from the bounces of each
objective to the bounce of an action in the scenario which preserve the sequential ordering.

@@

\modmf{%
The previous subsections permitted to define all the tools needed to build the Local Causality Graph (LCG).
This graph is denoted $\cwB$ where $u$ is the local state to be reached
from the initial state $s$.
It will be the basis of our static analysis,
as it represents the causality links between the different objectives involved
in the solving of the reachability of $u$.
This LCG is built recursively by considering some required local states (\eg $u$),
linking them to objectives (\eg $\PHobj{t}{u}$, if $t \in s$),
and locally refining these objectives in order to include new required local states
from other automata.
An example of LCG is depicted in \pref{fig:sa-livelock}.
}

Technically,
one can iteratively refine a given objective into (possibly several) objective
sequences extracted from its bounce sequences.
Indeed, the refinement of an objective $P$ with $\zeta\in\BS(P)$
consists in prepending to $P$ the
objectives leading to the activation of all hitters of the actions in $\zeta$,
in the same sequential order.
The generalization of this refinement to an objective sequence would
naturally require to consider all possible
interleaving between the refinements~\cite{PMR12-MSCS}.

@@

Formally, a Local Causality Graph (LCG)
is a digraph where $\cwV \subseteq \Proc\cup \Obj \cup \Sol \cup \sSol$ is the set of vertices,
with $\sSol=\powerset(\Proc)$ and $\Sol=\Obj\times\powerset(\sSol)$,
and $\cwE \subseteq \cwV \times \cwV$ is the set of oriented edges
(\pref{def:lcg}).
% 
% These refinements can be summarized with a so-called Local Causality Graph (LCG).
% Given an initial context $s$ and a local state $\myp\in\Proc$,
% an LCG $\cwB = (\cwV, \cwE)$
% is a digraph where $\cwV \subseteq \Proc\cup \Obj \cup \Sol \cup \sSol$ is the set of vertices
% where $\sSol=\powerset(\Proc)$ and $\Sol=\Obj\times\powerset(\sSol)$
% and $\cwE \subseteq \cwV \times \cwV$ is the set of oriented edges
% (\pref{def:lcg}).
A node in $\cwV\cap\Obj$ represents an objective to refine.
Such a node is linked to nodes in $\Sol$ which represent
sets of hitter sets extracted from a
bounce sequence of the objective
(that is, one node for each bounce sequence in $\BS(P)$); thus,
given an objective $P$, if $\zeta\in\BS(P)$
and $\abstr\zeta$ is the set of all hitters of $\zeta$,
% and $\abstr\zeta \DEF \{\PHhitter(\zeta_n) \mid
% n\in\indexes\zeta \}$,
then $P$ is linked to a node $\langle P, \abstr\zeta\rangle\in\Sol$
(\pref{def:lcg}\nobreakdash-\ref{lcg-obj-sol}).
For each hitter set $ps\in\abstr\zeta$,
the node $\langle P,\abstr\zeta\rangle\in\cwV\cap\Sol$ is
linked to the node $ps\in\sSol$
(\pref{def:lcg}\nobreakdash-\ref{lcg-sol-sync}).
Then, a node $ps\in\cwV\cap\sSol$ is linked to each local state $a_i \in ps$ is contains
(\pref{def:lcg}\nobreakdash-\ref{lcg-sync-ls}).
A local state node $a_i\in\cwV\cap\Proc$
is linked to each objective $\obj{a_j}{a_i} \in \Obj$, where
$a_j$ is either in the initial state $s$ or
is a local state node in the LCG
(\pref{def:lcg}\nobreakdash-\ref{lcg-ls-obj}).
Finally, given an objective node $\obj{a_j}{a_i}\in\cwV\cap\Obj$,
if a local state $a_k\neq a_i$ is in its successors (which is given by
$\conn_{(\cwBNodes,\cwBEdges)}(\obj{a_j}{a_i})$), then
$\obj{a_j}{a_i}$ is linked to $\obj{a_k}{a_i}$ in the LCG
(\pref{def:lcg}\nobreakdash-\ref{lcg-conn}).

\begin{definition}
\label{def:lcg}
Given a state $s \in \PHl$ and a local state $\myp\in \Proc$,
the Local Causality Graph (LCG) for reachability under-approximation
$\cwB \DEF (\Bv, \Be)$,
with
$\cwBNodes \subseteq \Proc \cup \Obj \cup \NSol \cup \sSol$
and
$\cwBEdges \subseteq \cwBNodes \times \cwBNodes$,
is the smallest graph such that:
\begin{enumerate}
\item
$\myp \in \cwBNodes$
\item
$a_i\in\cwBNodes\cap\Proc \Leftrightarrow \{ (a_i,\obj{a_j}{a_i}) \mid
a_j \neq \myp \wedge (a_j\in s \vee a_j\in\cwBNodes\cap\Proc)
\}\subseteq\cwBEdges$
\label{lcg-ls-obj}
\item
$P\in\cwBNodes\cap\Obj \Leftrightarrow 
	\{ (P,\langle P,\abstr \zeta\rangle) \mid \zeta\in\BS(P) \}\subseteq\cwBEdges$
\label{lcg-obj-sol}
\item
$\langle P, pps \rangle\in\cwBNodes\cap\NSol \Leftrightarrow
	\{ (\langle P,pps\rangle, ps) \mid ps\in pps \}\subseteq\cwBEdges$
\label{lcg-sol-sync}
\item
$ps \in\cwBNodes\cap\sSol \Leftrightarrow
	\{ (ps, a_i) \mid a_i\in ps \}\subseteq\cwBEdges$
\label{lcg-sync-ls}
\item
$\obj{a_i}{a_j}\in \cwBNodes\cap\Obj \Rightarrow 
	\{(\obj{a_i}{a_j},\obj{a_k}{a_j}) \mid a_k\neq a_j,$
\label{lcg-conn}
\\
\hspace*{4cm}
$a_k\in \conn_{(\cwBNodes,\cwBEdges)}(\langle\obj{a_i}{a_j},\abstr\zeta\rangle),$
\\
\hspace*{4cm}
$\zeta\in\BS(\obj{a_i}{a_j}) \}\subseteq\cwBEdges$
\end{enumerate}
with $\abstr\zeta \DEF \{\PHhitter(\zeta_n) \mid n\in\indexes\zeta \}$.
\end{definition}

\modmf{%
The definition of an LCG tackles two cases
where the target of an objective may be changed.
First,
the addition of objectives based on the local states already mentioned anywhere in the LCG,
as performed in \pref{def:lcg}\nobreakdash-\ref{lcg-ls-obj},
ensures to take into account the possible changes in the active local states
made by other objectives.
We thus try to ensure that a required local state is reachable even when
starting from a local state of the same automaton that is not in the initial state,
but that may be active at some point of the solving.
Second, \pref{def:lcg}\nobreakdash-\ref{lcg-conn},
allows to “re-target” objectives whose own solving changes their target.
This happens when playing the actions that activate the local states required to solve
an objective $P$ also changes the active local state of automaton $\PHsort(P)$.
In this case, the initial objective $P$ is re-targeted to another
objective $\PHobj{p}{\bounce{P}}$, where $p$ is the new active local state in $\PHsort(P)$.
Linking to all objectives of this kind ensures that all possible disturbances
are taken into account.
}

\modmf{%
Finally, we note that
the LCG of \pref{def:lcg} contains synchronizing nodes ($\sSol$)
that allow to express the need for several local states simultaneously
in order to play a given action.
This is the main difference regarding~\cite{PMR12-MSCS}
which tackled with Process Hitting, whose actions are limited to at most one hitter
and thus did not require this kind of synchronization.
However, \pref{def:lcg} is equivalent to the definition of the structure
given in~\cite{FPMR13-CS2Bio},
although it was adapted for AANs which notably allowed to simplify its formalism.
}

@@

Given a state $s \in \PHl$ and a local state $\myp\in\Proc$, the LCG $\cwB$ contains a set of objectives
that can be used to build a concrete scenario reaching $\myp$.
Because we are focused on sufficient conditions for reachability, we do not require that all
possible scenarios can be derived from the LCG.

In this section, we prove that if the LCG has no cycle, all its nodes in $\Obj$ have at least one
child, and all its nodes in $\sSol$ satisfy a
particular criteria, so-called independence,
then there exists a scenario that reaches $\myp$ from state $s$ (\pref{th:approxinf}).

@@

The intuition of the under-approximation of \pref{th:approxinf} is the following.
Given the initial state $s$, we recursively refine the initial objective
(which is: reaching $\myp$ from the initial state) according to its children.
As the LCG is acyclic, by hypothesis, such a recursion always terminates, and as all the objective nodes have at
least one child, it never gets stuck.
The refinement of an objective node $\obj{a_i}{a_j}$ acts as follows:
if the node has another objective node $\obj{a_k}{a_j}$ as child, we first refine the objective
$\obj{a_i}{a_k}$ (which, by construction, is necessarily in the LCG)
and then we refine the objective $\obj{a_k}{a_j}$.
If the objective node has only successors in $\Sol$ (bounce sequences), one is picked arbitrarily.
If $\langle P, pps\rangle$ is the chosen node,
by construction there exists $\zeta\in\BS(P)$ such that $\abstr\zeta = pps \in \powerset(\sSol)$.
If $\zeta = \emptyseq$ (for instance in the case where $a_i = a_j$), the recursion stops and we
continue to the next stage.
Otherwise, for each $n\in\indexes\zeta$,
for each $b_i \in\PHhitter(\zeta_n)$,
we refine the objective $\obj{b_j}{b_i}$, where $b_j$ is the state of $b$ in the current state.
By induction, $\obj{b_j}{b_i}$ is a child of $b_i$ in the LCG of \pref{def:lcg}.
Thus, we know that the current state of $b$ is $b_i$.
After having repeated this procedure for each $b_i\in\PHhitter(\zeta_n)$,
because all the nodes in $\sSol$, and in particular $\PHhitter(\zeta_n)$, are independent,
we know that all the local states in $\PHhitter(\zeta_n)$ are in the current state.
In addition, we know that the state of automaton $a$ has remained unchanged, otherwise
$\obj{a_i}{a_j}$ would have an objective child.
Hence, the action $\zeta_n$ is playable in the current state.
We can thus apply this action, modifying the state of $a$ to $a_j$ and continue to the next stage.
In the end, this recursive procedures builds a scenario from $s$ to a state containing $\myp$.

\begin{theorem}[Under-approximation]
\label{th:approxinf}
  Given an AAN $(\PHs; \PHl; \PHa)$,
  a state $s$ and local state $\myp$,
  if the LCG $\cwB$ contains no cycle,
  all nodes in $\Obj$ have at least one child,
  and all nodes in $\sSol$ are independent,
  then there exists a scenario $\delta\in\Sce$ such as $\myp\in s\play\delta$.
\end{theorem}

@@

We note furthermore that the method does not require any completeness of the bounce sequences
$\BS$.
Therefore, in order to reduce the number of bounce sequences to consider, and potentially remove
cycles and non-satisfying nodes,
one can consider only a sub-set of bounce sequences
(and in particular: a unique bounce sequence) for each objective.
However, such an approach requires to enumerate all possible combinations of bounce sequences
subsets, hence being exponential in the number of considered bounce sequences.
%objectives solved by at least two bounce sequences.

@@

    (b0) edge (b10) edge (b00)
    (b10) edge (b10s)
    (b10s) edge (nb10s)
    (b00) edge (b00s)
    (b00s) edge (nb00s)
    (b1) edge (b01) edge (b11)
    (b01) edge (b01s)
    (b01s) edge (b01ss)
    (b01ss) edge (a0)
    (b11) edge (b11s)
    (b11s) edge (nb11s)
    ;
    \end{tikzpicture}
  \caption{
  \label{fig:sa-livelock}
    The local causality graph $\cwB$ on the AAN in \pref{fig:ph-livelock}
    for the reachability of $\myp = c_1$
    from the initial state $s = \PHstate{a_1, b_0, c_0}$.
    Elements in $\Proc$ are represented by rectangular nodes,
    elements in $\Sol$ are represented by small circles,
    and elements in $\sSol$ and $\Obj$ are the remaining borderless nodes.
    \pref{th:approxinf} is inconclusive on this example as
    node $\{ a_1, b_1 \} \in \sSol$
    is not independent (see \pref{def:coherent}).
    Indeed, $a_0$ is a successor of $b_1$, but $a_0 \neq a_1$
    (and the same also stands for $b_0$, which is a successor of $a_1$).
  }
\end{figure}
\end{example}

@@

This section gives a recursive method to find a scenario that concretizes
the reachability of a given local state $\myp \in \Proc$,
from a given initial state $s_0 \in \PHl$,
provided that \pref{th:approxinf} answered positively on this couple of inputs.
The algorithm proposed in what follows relies on a traversal of the Local Causality Graph
$\cwBz$ that has been used for this conclusion.
The correctness of this extraction can be demonstrated
in the same fashion as \pref{th:approxinf}.

@@

The execution of such a traversal outputs a scenario from $s_0$.
In the following, we denote by $\Delta(\cwBz)$ the set of all scenarios extracted from the
local causality graph $\cwBz$.
Indeed, several scenarios may exist for a same graph,
generated by the arbitrary choices that exist in this algorithm.
An example of such a traversal is given on a small example
in \pref{ssec:ex-metazoan}.
